---
title: "Seurat and SpatialExperiment interaction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Seurat and SpatialExperiment interaction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE, warning=FALSE}
library(SpotGraphs)
library(Seurat)
library(TENxVisiumData)
library(SpatialExperiment)
library(dplyr)
library(igraph)
```

## Overview

In this vignette, we demonstrate how SpotGraphs can interact with Seurat and SpatialExperiment objects. A typical workflow will involve the following steps:

1. Extract x,y coordinates from Seurat or SpatialExperiment object.

2. Create igraph object with `SpotGraph()`.

3. Perform some analysis.

4. Store results back into the original Seurat or SpatialExperiment object.


## Seurat object interaction
We will use the example Seurat object provided in this package to demonstrate a typical workflow.
```{r, eval=TRUE}
class(scc_s1)
```

First, extract the x,y coordinates
```{r, eval=TRUE}
coord = Seurat::GetTissueCoordinates(scc_s1)
```

Create an igraph object with `SpotGraph()`
```{r, eval=TRUE}
ig = SpotGraph(coord[,c('x','y')])
```

Perform some analysis. In this case, we will calculate the number of edges per spot.
```{r, eval=TRUE}
n_edges = igraph::degree(ig)
```

Finally, store these results back into the metadata of the original Seurat object
```{r, eval=TRUE}
scc_s1 = AddMetaData(scc_s1, metadata = n_edges, col.name = 'degree')
```

## SpatialExperiment interaction
To demonstrate interaction with a SpatialExperiment object, we will load an object from ExperimentHub. 
```{r, eval=TRUE, message=FALSE}
eh = ExperimentHub::ExperimentHub()
q = AnnotationHub::query(eh, 'TENxVisium')
id = q$ah_id[q$title=='HumanGlioblastoma_v3.13']
spe = eh[[id]]
class(spe)
```

Once we have our SpatialExperiment object, we can follow the same workflow
```{r, eval=TRUE}
coord = as.data.frame(spatialCoords(spe))
ig = SpotGraph(coord)
n_edges = igraph::degree(ig)
colData(spe)$degree = n_edges
```

We could also calculate multiple statistics from several analyses on the igraph object before storing them in the original Seurat/SpatialExperiment object. If we store all the vertex/spot-level results back into the igraph object with `V()`, we can extract all of the results at once with `igraph::vertex_attr()`, and coerce this into a `data.frame`.
```{r, eval=TRUE}
V(ig)$cluster = factor(cluster_louvain(ig)$membership)
V(ig)$ecc = eccentricity(ig)
V(ig)$centrality = centr_eigen(ig)$vector

df = as.data.frame(vertex_attr(ig))
head(df)
```

This resulting `data.frame` and then be stored back into the SpatialExperiment object. 
```{r}
colData(spe) = cbind(colData(spe), df)
head(colData(spe))
```



