---
title: "Identify centers or boundaries"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Identify centers or boundaries}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE, eval=TRUE}
library(SpotGraphs)
library(Seurat)
library(ggplot2)
library(dplyr)
library(viridis)
library(patchwork)
```

# Overview
A common task in spatial analysis is to perform some analysis with a region of interest, such as a tumor. In this vignette, we demonstrate how to use the `NeighborhoodCenters()` function to help us identify characteristics such as the center of a region of interest or its boundary. 

# 1. Load data
We will again use the 10X Visium dataset, downloaded from GSE208253, to demonstrate `NeighborhoodCenters()`.

```{r Load data}
data("scc_s1", package = "SpotGraphs")
scc_s1 = UpdateSeuratObject(scc_s1)
class(scc_s1)
dim(scc_s1)

coord = Seurat::GetTissueCoordinates(scc_s1)
coord = data.frame(x = coord$y, y = -coord$x)
head(coord)
```

# 2. Select a region of interest
In this vignette, we will use clustering to select some region of interest in this tissue sample, but in practice, this should instead be a region of some biological interest. The region of interest, i.e., neighborhood, is provided to `NeighborhoodCenters()` as a named vector of `TRUE`/`FALSE` values, where the names correspond to spot barcodes in the given dataset.
```{r Choose a region}
# create an igraph object with SpotGraph() and perform clustering
ig = SpotGraph(coord)
cl = igraph::cluster_fast_greedy(ig)
cl = factor(cl$membership)

# select cluster 1 and name the boolean vector with vertex names
roi = cl==5
names(roi) = names(igraph::V(ig))
```

# 3. Run `NeighborhoodCenters()`
```{r Run NeighborhoodCenters}
res = NeighborhoodCenters(coord = coord, is_neighborhood = roi)
```

The output from this function returns a list, where the first element is a data.frame with the scores from `igraph::centr_eigen()` where `is_neighborhood == T`, and 0 where `is_neighborhood == F`. 
```{r Plot eigenvector centrality scores, fig.width=7.75, fig.height=7.5}
print(head(res$eigen.scores))

# Set centrality scores as a vertex attribute in our igraph object
igraph::V(ig)$centr_eigen = res$eigen.scores$center_eigen
igraph::V(ig)$roi_center = names(igraph::V(ig)) %in% res$centers
igraph::V(ig)$roi_boundary = names(igraph::V(ig)) %in% res$boundary

plt1 = SpatialPlotGraph(ig, group.by = 'centr_eigen',
                        pt.size = 0.5, linewidth = 0.1) +
  scale_color_viridis() +
  theme_void()
plt2 = SpatialPlotGraph(ig, group.by = 'roi_center',
                        pt.size = 0.5, linewidth = 0.1) +
  scale_color_manual(values = c('grey', 'red')) +
  theme_void() +
  theme(legend.position = 'none')
plt3 = SpatialPlotGraph(ig, group.by = 'roi_boundary', 
                        pt.size = 0.5, linewidth = 0.1) +
  scale_color_manual(values = c('grey', 'red')) +
  theme_void()

wrap_plots(plt1, plt2, plt3, design = 'A#\nBC')
```

# 4. Select multiple regions of interest
The `NeighborhoodCenters()` function can also identify characteristics of multiple regions by providing a named boolean vector indicating whether a spot is within any of these regions. This can be useful when there are several tumor regions within the same tissue section. Here, we select three different clusters as regions of interest, instead of just one.

```{r NeighborhoodCenters with multiple regions}
roi = cl %in% c('1', '2', '5')
names(roi) = names(igraph::V(ig))
res = NeighborhoodCenters(coord = coord, is_neighborhood = roi)
```

```{r multiple region plotting, fig.width=7.75, fig.height=7.5}
# Set centrality scores as a vertex attribute in our igraph object
igraph::V(ig)$centr_eigen = res$eigen.scores$center_eigen
igraph::V(ig)$roi_center = names(igraph::V(ig)) %in% res$centers
igraph::V(ig)$roi_boundary = names(igraph::V(ig)) %in% res$boundary

plt1 = SpatialPlotGraph(ig, group.by = 'centr_eigen', 
                        pt.size = 0.5, linewidth = 0.1) +
  scale_color_viridis() +
  theme_void()
plt2 = SpatialPlotGraph(ig, group.by = 'roi_center', 
                        pt.size = 0.5, linewidth = 0.1) +
  scale_color_manual(values = c('grey', 'red')) +
  theme_void() +
  theme(legend.position = 'none')
plt3 = SpatialPlotGraph(ig, group.by = 'roi_boundary', 
                        pt.size = 0.5, linewidth = 0.1) +
  scale_color_manual(values = c('grey', 'red')) +
  theme_void()

wrap_plots(plt1, plt2, plt3, design = 'A#\nBC')
```

# 5. Calculate distance between each spot and ROI center
```{r ROI distances, fig.width=8, fig.height=4.75}
roi_dist = igraph::distances(ig, weights = NA)[,res$centers] %>% 
  apply(1, min)
boundary_dist = igraph::distances(ig, weights = NA)[,res$boundary] %>% 
  apply(1, min)
# boundary_dist = ifelse(!roi, boundary_dist, 0)

igraph::V(ig)$roi_dist = roi_dist
igraph::V(ig)$boundary_dist = boundary_dist

plt.dist = SpatialPlotGraph(ig, group.by = 'roi_dist',
                 pt.size = 0.5, linewidth = 0.1) +
  scale_color_distiller(palette = 'RdBu', direction = 1) +
  ggtitle('distance to ROI centers') +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5, face = 'bold'),
        legend.position = 'none')
plt.boundarydist = SpatialPlotGraph(ig, group.by = 'boundary_dist',
                 pt.size = 0.5, linewidth = 0.1) +
  scale_color_distiller(palette = 'RdBu', direction = 1) +
  ggtitle('distance to ROI boundaries') +
  guides(color = guide_colorbar(title = 'distance'),
         alpha = guide_none()) +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5, face = 'bold'))

wrap_plots(plt.dist, plt.boundarydist, nrow = 1)
```


