---
title: "SpotGraphs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SpotGraphs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE, warning=FALSE}
library(SpotGraphs)
library(TENxVisiumData)
library(dplyr)
library(ggplot2)
library(viridis)
library(ggnetwork)
```

## Load example dataset
To demonstrate the general usage of SpotGraph, will use a 10X Visium dataset from a squamous cell carcinoma patient, downloaded from GSE208253. The raw data from GEO is provided as a Seurat object in this package to use as an example.


```{r}
class(scc_s1)
dim(scc_s1)
```

## Create an igraph object
Our goal with this package is support Visium analysis by providing a function that constructs an igraph object based on the number of immediately adjacent spots on a slide. This opens up the analysis pipeline to various igraph functions to perform certain tasks such as:

1. identifying spots lying on tissue debris to be removed from downstream analysis
2. distance calculations between spots by measuring the shortest path through the network between each spot
3. easily identifying edges of pre-defined regions of tissue

To construct an igraph object, all we need are the x,y coordinates of the spots on our slide.

```{r}
coord = Seurat::GetTissueCoordinates(scc_s1)
coord = coord[,c('x', 'y')]
head(coord)

ig = SpotGraph(coord = coord)
ig
```

We can plot this igraph object with the ggnetwork package, which interfaces directly with ggplot2, or use the SpatialPlotGraph function provided in this package to visualize the edges drawn between each node in the original x,y coordinates stored in the igraph object. Note that the `is_boundary` attribute is automatically calculated from running `SpotGraph()`, which we will use to color each node in our plots.

```{r, fig.width=9, fig.height=4}
plt.ggnet = ggplot(ig, aes(x=x, y=y, xend=xend, yend=yend)) +
  geom_edges() + 
  geom_nodes(aes(color = is_boundary), size = 0.5) +
  theme_void() +
  theme(legend.position = 'none')

plt.spg = SpatialPlotGraph(igraph_object = ig, 
                           group.by = 'is_boundary', 
                           flip.axes = T, 
                           pt.size = 0.5)

patchwork::wrap_plots(plt.ggnet, plt.spg)
```


## Spot filtering
One of the main features of the SpotGraph package is to identify spots on a slide that lie on top of tissue debris disconnected from the rest of the tissue sample. These are uninformative and are a technical artifact from sample processing, so we'd like to exclude these from our downstream analysis. The `CleanSlide()` function provides a simple way to automatically identify these spots, based on connectivity to other spots and the number of transcripts detected within each community of spots. 

First, we extract the two inputs required for the `CleanSlide()` function:

1. the x,y coordinates of each spot in our dataset
2. a vector of transcript counts detected in each spot
```{r}
coord = Seurat::GetTissueCoordinates(scc_s1)
coord = coord[,c('x', 'y')]
nCounts = scc_s1$nCount_Spatial
```

The output of `CleanSlide()` is a data.frame with the same number of rows as the input coordinate data.frame or matrix that can be directly used with `Seurat::AddMetaData()` to re-insert back into the original Seurat object.
```{r}
res = CleanSlide(coord, nCount = nCounts)
scc_s1 = Seurat::AddMetaData(scc_s1, res)
```

We can now observe the results of `CleanSlide()` with `Seurat::SpatialDimPlot()`
```{r, fig.width=6, fig.height=4, message=FALSE}
Seurat::SpatialDimPlot(scc_s1, group.by = 'threshold', image.alpha = 0.6)
```

We can agree with these results and directly apply the threshold to filter out these spots with `subset(scc_s1, threshold)`, or we can manually identify groups of spots that we want to filter out. The 'ig_cluster' metadata column stores communities of spots, where each spot is immediately adjacent to other spots. 
```{r, fig.width=6, fig.height=4, message=FALSE, eval=T}
Seurat::SpatialDimPlot(scc_s1, group.by = 'ig_cluster', image.alpha = 0.6, label = T, label.size = 3)
```

In this case, we could manually remove clusters 9, 10, 23, 24, 26, 22, 15, 18, 19, 17, 20, and 25. 
```{r, fig.width=6, fig.height=4, message=FALSE, warning=FALSE, eval=T}
spot_clusters_to_remove = c(9, 10, 23, 24, 26, 22, 15, 18, 19, 17, 20, 25)
scc_s1 = scc_s1[, !scc_s1$ig_cluster %in% spot_clusters_to_remove]
Seurat::SpatialFeaturePlot(scc_s1, features = 'nCount_Spatial', pt.size.factor = 2)
```

## CutEdges
When working with an igraph object, it may be desirable to remove all edges between two groups of spots. 

```{r, fig.width=11, fig.height=4, eval=T}
# Create igraph object
coord = Seurat::GetTissueCoordinates(scc_s1)
coord = coord[,c('x', 'y')]
ig = SpotGraph(coord)

# Clustering with igraph
igraph::V(ig)$clusters = factor(igraph::cluster_fast_greedy(ig)$membership)
plt1 = SpatialPlotGraph(ig, group.by = 'clusters', label = T) +
  ggtitle('before') +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5, face = 'bold'),
        legend.position = 'none')

# Remove edges between pairs of clusters
ig = CutEdges(ig, cluster_pairs = list(c(2,1), c(2,10)), cluster.col = 'clusters')
plt2 = SpatialPlotGraph(ig, group.by = 'clusters', label = T) +
  ggtitle('after') +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5, face = 'bold'),
        legend.position = 'none')

plt3 = plt2 + 
  coord_cartesian(xlim = c(0.35,0.8), ylim = c(0.4, 1)) +
  ggtitle('after/zoomed')
patchwork::wrap_plots(plt1, plt2, plt3)
```

