---
title: "SpotGraphs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SpotGraphs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE, warning=FALSE}
library(SpotGraphs)
library(Seurat)
library(dplyr)
library(ggplot2)
library(patchwork)
library(viridis)
library(ggnetwork)
```

## Load example dataset
To demonstrate the general usage of SpotGraph, will use a 10X Visium dataset from a squamous cell carcinoma patient, downloaded from GSE208253. The raw data from GEO is provided as a Seurat object in this package to use as an example.


```{r, message=FALSE}
data("scc_s1", package = "SpotGraphs")
scc_s1 = UpdateSeuratObject(scc_s1)
class(scc_s1)
dim(scc_s1)
```

## Create an igraph object
Our goal with this package is support Visium analysis by providing a function that constructs an igraph object based on the number of immediately adjacent spots on a slide. This opens up the analysis pipeline to various igraph functions to perform certain tasks such as:

1. identifying spots lying on tissue debris to be removed from downstream analysis
2. distance calculations between spots by measuring the shortest path through the network between each spot
3. easily identifying edges of pre-defined regions of tissue

To construct an igraph object, all we need are the x,y coordinates of the spots on our slide.

```{r}
coord = Seurat::GetTissueCoordinates(scc_s1)
coord = data.frame(x = coord$y, y = -coord$x)
head(coord)

ig = SpotGraph(coord = coord)
ig
```

We can plot this igraph object with the ggnetwork package, which interfaces directly with ggplot2, or use the SpatialPlotGraph function provided in this package to visualize the edges drawn between each node in the original x,y coordinates stored in the igraph object. Note that the `is_boundary` attribute is automatically calculated from running `SpotGraph()`, which we will use to color each node in our plots.

```{r, fig.width=9, fig.height=4}
plt.ggnet = ggplot(ig, aes(x=x, y=y, xend=xend, yend=yend)) +
  geom_edges() + 
  geom_nodes(aes(color = is_boundary), size = 0.5) +
  theme_void() +
  theme(legend.position = 'none')

plt.spg = SpatialPlotGraph(igraph_object = ig, 
                           group.by = 'is_boundary', 
                           pt.size = 0.5)

patchwork::wrap_plots(plt.ggnet, plt.spg)
```


## Spot filtering based on connectivity
One of the main features of the SpotGraph package is to identify spots on a slide that lie on top of tissue debris disconnected from the rest of the tissue sample. These are uninformative and are a technical artifacts from sample processing, so we'd like to exclude these from our downstream analysis. We can identify whether a spot is surrounded by other spots using the `igraph::degree()` function, and filter out spots that are by themselves or only adjacent to one other spot with `igraph::subgraph()`:

1. First, extract the x,y coordinates of each spot in our dataset
2. Create and igraph object with `SpotGraph()`
3. Calculate the number of connections per-spot with `igraph::degree()`
```{r}
coord = GetTissueCoordinates(scc_s1)
ig = SpotGraph(coord[,1:2])

n_connections = igraph::degree(ig)
```

We can now plot the number of connections (i.e., the results of `igraph::degree()`) by storing these results back into the original Seurat object that we started with. Additionally, we can add another metadata column to indicate whether each spot is adjacent to more than one other spot. 
```{r}
scc_s1 = AddMetaData(scc_s1, n_connections, 'degree')
scc_s1 = AddMetaData(scc_s1, n_connections>1, 'adj_threshold')
```

We can now observe the results of `igraph::degree()` with `Seurat::SpatialFeaturePlot()`. Spots colored in red are the spots that we've identified to only have one adjacent spot or are by themselves on our slide. With this method, we would be able to further remove low quality spots from our data, entirely based on spot-level adjacencies. 
```{r, fig.width=9, fig.height=4, message=FALSE}
plt.deg = SpatialFeaturePlot(scc_s1, feature = 'degree', image.alpha = 0.6) +
  scale_fill_distiller(palette = 'Blues')
plt.degfilter = SpatialDimPlot(scc_s1, group.by = 'adj_threshold', image.alpha = 0.6) +
  scale_fill_manual(values = c('red', 'grey90'))
wrap_plots(plt.deg, plt.degfilter)
```

Alternatively, we could perform clustering using the spot adjacencies we identified with `SpotGraph()`. We set the `resolution = 0` to group all spots together that are connected in any way. We can store these clustering results back into our Seurat object and again, plot with `Seurat::SpatialDimPlot()`. After clustering, we can choose to keep only clusters 1 and 2, which will let us remove any low quality spots that are disconnected from most of the tissue sample in our data.
```{r, fig.width=9, fig.height=4, message=FALSE, eval=T}
cl = igraph::cluster_louvain(ig, resolution = 0)
scc_s1 = AddMetaData(scc_s1, factor(cl$membership), 'igraph_clusters')
plt.cl = SpatialDimPlot(
  object = scc_s1, 
  group.by = 'igraph_clusters', 
  image.alpha = 0.6, 
  label = T, 
  label.size = 3
) + NoLegend()
  

scc_s1 = AddMetaData(scc_s1, cl$membership %in% c('1','2'), 'cl_threshold')
plt.clfitler = SpatialDimPlot(
  object = scc_s1, 
  group.by = 'cl_threshold', 
  image.alpha = 0.6
) + scale_fill_manual(values = c('red', 'grey90'))

wrap_plots(plt.cl, plt.clfitler)
```

We will apply the filter determined from using the igraph clustering method. 
```{r, message = FALSE, warning=FALSE}
# Apply filter to keep igraph clusters 1 and 2
scc_s1 = subset(scc_s1, subset = cl_threshold)
```


## CutEdges
When working with an igraph object, it may be desirable to remove all edges between two groups of spots. 

```{r, fig.width=11, fig.height=4, eval=T}
# Create igraph object
coord = GetTissueCoordinates(scc_s1)
# coord = data.frame(x = coord$y, y = -coord$x)
coord = coord[,c('x', 'y')]
ig = SpotGraph(coord)

# Clustering with igraph
igraph::V(ig)$clusters = factor(igraph::cluster_fast_greedy(ig)$membership)
plt1 = SpatialPlotGraph(ig, group.by = 'clusters', label = T) +
  ggtitle('before') +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5, face = 'bold'),
        legend.position = 'none')

# Remove edges between pairs of clusters
ig = CutEdges(ig, cluster_pairs = list(c(4,7), c(4,3), c(4,8)), cluster.col = 'clusters')
plt2 = SpatialPlotGraph(ig, group.by = 'clusters', label = T) +
  ggtitle('after') +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5, face = 'bold'),
        legend.position = 'none')

plt3 = plt2 + 
  coord_cartesian(xlim = c(0.15,0.65), ylim = c(0.15, 0.65)) +
  ggtitle('after/zoomed')
wrap_plots(plt1, plt2, plt3)
```

