---
title: "ThresholdTests"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ThresholdTests}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Test different strategies to automatically detect a threshold for filtering

```{r setup}
library(SpotGraphs)
library(dplyr)
library(stringr)
library(tidyr)
library(ggplot2)
library(viridis)
library(patchwork)
```

## Load several datasets to get a survey of potential cluster_nCount distributions
```{r}
# retrieve data from TENxVisium
eh = ExperimentHub::ExperimentHub()
q = AnnotationHub::query(eh, 'TENxVisium')
ds.names = c('HumanLymphNode', 
             'MouseBrainCoronal',
             'HumanColorectalCancer_v3.13',
             'HumanSpinalCord_v3.13',
             'HumanGlioblastoma_v3.13',
             'MouseKidneyCoronal_v3.13')
id = q$ah_id[q$title %in% ds.names]
spe.ls = lapply(id, function(ds.id) eh[[ds.id]])
names(spe.ls) = ds.names %>% str_replace('\\.', '_')
```

## Use CleanSlide to get cluster_nCounts
```{r}
spe.ls = lapply(spe.ls, function(spe) {
  colData(spe)$nCount = colSums(assay(spe))
  return(spe)
})
res.ls = lapply(spe.ls, function(spe) {
  coord = SpatialExperiment::spatialCoords(spe) %>% as.data.frame()
  colnames(coord) = c('x', 'y')
  CleanSlide(coord = coord, nCount = spe$nCount)
})
```

## Store the result dataframe in each SpatialExperiment object
```{r}
for (ds.name in names(res.ls)) {
  colData(spe.ls[[ds.name]]) = cbind(colData(spe.ls[[ds.name]]), res.ls[[ds.name]])
}
```


## Plot distributions of each dataset
```{r}
den = density(log10(cluster_nCount.ls$HumanLymphNode))
df = data.frame(x=den$x, y=den$y)
ggplot(df, aes(x=x, y=y)) +
  geom_line() +
  theme_classic()
```



```{r}
ImageFeature = function(obj, feature = 'nCount', label = F) {
  # First, scale the x,y coordinates to fit the low-res H&E image
  lowres_scale <- imgData(obj)[imgData(obj)$image_id == 'lowres', 'scaleFactor']
  coord.xy = SpatialExperiment::spatialCoords(obj) %>% as.data.frame()
  coord.xy$x_axis <- coord.xy$pxl_row_in_fullres * lowres_scale
  coord.xy$y_axis <- coord.xy$pxl_col_in_fullres * lowres_scale
    
  # Flip the y-axis
  coord.xy$y_axis <- abs(coord.xy$y_axis - (ncol(imgRaster(obj)) + 1))
    
  # Get metadata feature from object
  coord.xy[,feature] = colData(obj)[,feature]

  # Create plot of transcripts per spot overlaid ontop of H&E
  plt = ggplot(mapping = aes(1:600, 1:600)) +
    annotation_raster(imgRaster(obj), 
                      xmin = 1, xmax = 600, 
                      ymin = 1, ymax = 600) +
    geom_point(data=coord.xy, alpha=0.5, size = 0.5,
               aes(x=x_axis, y=y_axis, colour=get(feature))) +
    scale_color_viridis(option = 'turbo', name = feature) +
    coord_cartesian(xlim = c(1,600), 
                    ylim = c(1,600)) +
    coord_fixed() +
    theme_void()
  
  if (label) {
    df.labels = coord.xy %>% 
      reframe(.by = feature, 
              x_axis = mean(x_axis),
              y_axis = mean(y_axis))
    
    plt = plt +
      geom_label(data = df.labels, aes(x=x_axis, y=y_axis, label = get(feature)))
  }
  return(plt)
}

tmp = ImageFeature(spe.ls[[1]], feature = 'cluster_nCount', label = T)
plt.ncount = lapply(names(spe.ls), function(ds.name) {
  ImageFeature(spe.ls[[ds.name]], feature = 'cluster_nCount', label = T) +
    ggtitle(ds.name) +
    theme(plot.title = element_text(hjust = 0.5, face = 'bold'))
})
wrap_plots(plt.ncount, ncol = 3)
```

## Extract cluster_nCount per dataset
```{r}
cluster_nCount.ls = lapply(res.ls, function(res) {
  unique(res$cluster_nCount)
})
```

## Write the current helper function to identify the threshold
```{r}
# helper to calculate threshold, given a vector of transcript counts
get_threshold = function(cluster_nCount) {
  # log the total counts per cluster to smooth out the density
  den = density(log10(cluster_nCount))

  # define an interval to check for local minima
  int.min = quantile(den$x, 0.05) #previous value of 0 might be to low
  int.max = quantile(den$x, 0.75) #maybe 3rd quantile?
  interval = c(int.min, int.max)

  # calculate local minima
  thres = approxfun(den$x, den$y) %>% optimise(interval=interval)
  thres = 10^(thres$minimum)
  return(thres)
}
```

